---
title: "Very basic R Notebook intro"
output:
 html_document
---

If you open an Rnotebook in RStudio, it is going to give a similar intro as this first part:

Everything that is written outside r chunks (first ones comes below), is interpreted in Markdown language.

Very quick and short intro to RMarkdown:

# This is a first level section header
Very general stuff.
## This is a second level section header
Related content that deserves its own subsection.
### This is a third level section header
Related content that deserves its own subsubsection.
## This is another second level section header
More related content that deserves its own subsection.

<!-- This is how you comment things in RMarkdown. They won't show up in the final odf/html file -->
<!-- But note that commented code inside visible chunks will -->

For now, for further information, check:
- [bookdown website](https://bookdown.org/)
- [How to write reproducible manuscripts](https://elifesciences.org/labs/cad57bcf/composing-reproducible-manuscripts-using-r-markdown)

# R chunks

This a very simple R chunk
```{r}
iris
```

This a named R chunk. The name of the r chunk does not appear in the pdf/html, but it might be useful for internal code organization and creating cross-references (more on this later). 
```{r check-iris}
iris
```

R chunks have a few options regarding its visibility (by default, they are all `TRUE` and can be set up in the UI of RStudio):
- `eval=FALSE`: none of the code in the chunk is evaluated when compiling the pdf/html
- `include=FALSE`: the chunk does not appear in the pdf/html
- `echo=FALSE`: the outputs of the chunk are not shown in the pdf/html
- `warnings=FALSE`: warnings generated by code in the chunk are not shown
- `messages=FALSE`: messages generated by the code in the chunk are not shown
