---
title: "Example of a computational notebook for data analysis"
output:
  html_document:
    df_print: paged
    code_folding: hide
  pdf_document: default
---

```{r set-up, include = FALSE}
## This chunk of code is not relevant for understanding the work done in this document.
## Thus, it won't appear in the pdf or html 
rawdata_dir <- file.path("results", "data", "raw") ## do NOT play with stuff in raw_data. That is your back-up
scripts_dir <- file.path("results", "scripts")
procdata_dir <- file.path("results", "data", "processed")
figures_dir <- file.path("text", "figures")
tables_dir <- file.path("text", "tables")
suppl_dir <- file.path("text", "supplementary")

# internal formatting of the notebook file
knitr::opts_chunk$set(message = FALSE)

# custom aesthetics
options(ggplot2.continuous.colour = pals::ocean.oxy(n = 27))
options(ggplot2.discrete.colour = pals::ocean.oxy(n = 27))

## base theme for plotting figures
theme_lemurs <- function(){
    theme_minimal(base_size = 8)%+replace%
        theme(
            axis.text = element_text(size = rel(1.25)),
            axis.title.x = element_text(size = rel(1.25)),
            axis.title.y = element_text(size = rel(1.25), angle = 90),
            strip.text = element_text(size = 10, face = "bold"),
            strip.background = element_rect(colour = "#E6E6F4",
                                            fill = "#E6E6F4"),
            legend.title = element_text(face = "bold", size = 10),
            legend.text = element_text(size = 10),
            legend.position="bottom",
            legend.background = element_rect(fill = "transparent",
                                             colour = NA),
            legend.key = element_rect(fill="transparent", colour=NA)
    )
}

```

```{r package-upload, echo=FALSE,message=FALSE}
library(tidyverse) ## data processing
library(ggplot2) ##visualization
library(ggridges)
library(kableExtra) ## nice looking tables for html and pdf files. Used here with default configuration. Simply calling the data frames or their header would have a similar, if less aesthetically pleasing result. 
```
In this example, we explore life history data on 2270 lemur individuals living in the Duke Lemur Center. 

# Data

In this example, we explore life history data on 2270 lemur individuals living in the [Duke Lemur Center](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-24/readme.md).

```{r}
lemurs_rawdf <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv',
                          col_types = cols(
                            .default = col_double(),
                            taxon = col_character(),
                            dlc_id = col_character(),
                            hybrid = col_character(),
                            sex = col_character(),
                            name = col_character(),
                            current_resident = col_character(),
                            stud_book = col_character(),
                            dob = col_date(format = ""),
                            estimated_dob = col_character(),
                            birth_type = col_character(),
                            birth_institution = col_character(),
                            estimated_concep = col_date(format = ""),
                            dam_id = col_character(),
                            dam_name = col_character(),
                            dam_taxon = col_character(),
                            dam_dob = col_date(format = ""),
                            sire_id = col_character(),
                            sire_name = col_character(),
                            sire_taxon = col_character(),
                            sire_dob = col_date(format = ""),
                            dod = col_date(format = ""),
                            age_of_living_y = col_double(), ## this column is typed wrong (as character) by default
                            dob_estimated = col_character(),
                            weight_date = col_date(format = ""),
                            age_category = col_character(),
                            preg_status = col_character(),
                            concep_date_if_preg = col_date(format = ""),
                            infant_dob_if_preg = col_date(format = "")
                            )
)

## key of species name and abbreviation in the taxon columns
lemurs_sppnames_df <- readr::read_csv(file.path(rawdata_dir, "lemurs_sppnames.csv"), 
                                      col_names = TRUE)

lemurs_rawdf %>%
  head() %>%
  kableExtra::kbl(caption = "Header of the original data set on the health, reproduction, and social dynamics of lemurs housed at the Duke Lemur Center, in North Carolina, USA.")%>%
  kableExtra::kable_styling(c("striped", "hover")) %>%
  kableExtra::scroll_box(width = "100%", height = "300px")
```

## Pre-processing

The data can be organized temporally, thanks to the `weight_date` and `month_of_weight` variables, which report the full date and the month when the weight was measured, respectively. 
Moreover, we do not need all the 52 variables that were measured, so let's create smaller time-series with the variables of interest.
```{r}
lemurs_smallts <- lemurs_rawdf %>%
  dplyr::mutate_at(vars(name, dam_name, sire_name), stringr::str_to_title) %>%
  dplyr::mutate(year = lubridate::year(weight_date)) %>%
  dplyr::select(c(year, month_of_weight, ## time variables
                  taxon, dlc_id, ## id variables
                  hybrid, sex, name, birth_month, litter_size, concep_month, ## birth variables
                  dam_id, dam_name, dam_name, sire_id, sire_name, sire_taxon, ## parental history variables
                  age_at_death_y, age_of_living_y, age_last_verified_y, 
                  age_max_live_or_dead_y, age_at_wt_y, age_category, ## age variables
                  weight_g, avg_daily_wt_change_g, ## weight variables
                  preg_status, 
                  n_known_offspring, infant_lit_sz_if_preg)) %>%
  dplyr::rename(month = month_of_weight,
                weight = weight_g,
                avg_d_wt_chg = avg_daily_wt_change_g,
                n_offspring = n_known_offspring) %>%
  dplyr::right_join(lemurs_sppnames_df,., by = "taxon") ## id species
```

# Main text figures and tables

## Table 1: Females fertility rates per taxon

```{r}
lemurs_smallts %>%
  dplyr::filter(!is.na(infant_lit_sz_if_preg)) %>% ## filter the animals for which this information was available
  dplyr::group_by(dlc_id, species) %>%
  dplyr::summarize(inflt_mean_ind = mean(infant_lit_sz_if_preg)) %>%
  ungroup() %>%
  dplyr::group_by(species) %>%
  dplyr::summarize(inflt_mean = mean(inflt_mean_ind),
                   inflt_sd = sd(inflt_mean_ind),
                   n = n()) %>%
  dplyr::rename(Species = species,
                "Infant litter size (mean)" = inflt_mean,
                "Infant litter size (sd)" = inflt_sd) %>%
  readr::write_csv(file.path(figures_dir, "fertility_rates.csv"))
```

## Seasonality of species
```{r}
births_df <- lemurs_smallts %>%
  dplyr::select(dlc_id, taxon, birth_month, year) %>%
  dplyr::filter(!is.na(birth_month)) %>%
  dplyr::mutate_at(vars(birth_month), 
                   lubridate::month, label = TRUE, 
                   locale = Sys.getlocale(category = "LC_CTYPE")) %>% ## id months
  dplyr::right_join(lemurs_sppnames_df,., by = "taxon") %>% ## id species
  dplyr::arrange(taxon, birth_month) 
 
months_fct <- lubridate::month(1:12, label = TRUE, locale = Sys.getlocale(category = "LC_CTYPE"))

births_countdf <- births_df %>%
  unique() %>%
  dplyr::group_by(species, common_name, taxon, birth_month, year) %>%
  dplyr::summarize(n_births = n()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(species, common_name, taxon, year), 
                     names_from = birth_month, values_from = n_births) %>%
  tidyr::pivot_longer(all_of(months_fct), 
                      names_to = "birth_month", values_to = "n_births")%>%
  ## because strings can't be converted back to months, the following turns out quite cumbersome
  dplyr::mutate_at(vars(birth_month), 
                   ~ ordered(.,
                             levels = all_of(months_fct)) %>%
                     as.numeric) %>%
  dplyr::mutate_at(vars(birth_month), 
                   lubridate::month, label = TRUE,
                                       locale = Sys.getlocale(category = "LC_CTYPE")) %>%
  dplyr::arrange(year, species, birth_month) 

births_summdf <- births_countdf  %>%
  dplyr::group_by(species, birth_month) %>%
  dplyr::summarize(n_births_mean = mean(n_births, na.rm = TRUE),
                 n_births_sd = sd(n_births, na.rm = TRUE))
```

```{r fig.height=6, fig.width=12, fig.caption = "Mean number of individuals born in each month of the year, averaged for species across years (1938-2019)."}
births_summdf %>%
  ggplot(aes(x = birth_month, y = n_births_mean, colour = species, group = species)) +
  geom_step() +
  theme_lemurs() + 
  labs(x = "Month", y = "Number of births") +
  theme(legend.position = "bottom")
```

```{r fig.height=10, fig.width=15}
births_summdf  %>%
  ggplot(aes(x = birth_month, y = n_births_mean, fill = species)) +
  geom_bar(alpha=0.6, stat = "identity") + 
  theme_lemurs() + 
  facet_wrap(~species, ncol = 3) +
  labs(x = "Month", y = "Number of births (mean)") +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45))
```

## Offspring production
```{r}
offspring_df <- lemurs_smallts %>%
  dplyr::select(year, month, species, taxon, dlc_id, sex,
                litter_size, ## size of litter it was born into
                age_at_wt_y, weight, 
                preg_status, 
                n_offspring, ## total number of offspring produced until that day
                infant_lit_sz_if_preg)
```

## Individual weight and litter size
Get individual's weight at its younger age and plot it against against the litter it came from (separate males and females differently)
```{r}
litterweight_df <- offspring_df %>%
  dplyr::group_by(dlc_id) %>%
  dplyr::filter(age_at_wt_y == min(age_at_wt_y)) %>%
  ungroup()

litterweight_df %>%
  dplyr::group_by(species) %>%
  dplyr::summarize(weight_mean = mean(weight),
                   weight_sd = sd(weight))
```

```{r}
litterweight_df  %>%
  dplyr::filter(sex != "ND") %>%
  ggplot(aes(x = log(weight), y = species))+
  geom_point(aes(size = litter_size), alpha = 0.2) +
  facet_wrap(~sex, ncol = 2) +
  labs(x = "Weight (log(g))", y = "Species", size = "Litter size") + 
  theme_lemurs()
```

## Individual female weight and offspring production
```{r}
offspring_df  %>%
  dplyr::filter(preg_status == "P") %>%
  dplyr::group_by(dlc_id, species) %>%
  ggplot(aes(x = log(weight), y = species))+
  geom_point(aes(size = infant_lit_sz_if_preg), alpha = 0.2) +
  labs(x = "Weight (log(g))", y = "Species", size = "Infant litter size") +
  theme_lemurs()
```

TODO: get weight of each individual at its oldest, and plot it against the n_offspring it produced. Facet for males and females
```{r}
offspring_df  %>%
  filter(sex != "ND") %>%
  dplyr::group_by(dlc_id) %>%
  dplyr::filter(age_at_wt_y == max(age_at_wt_y)) %>%
  ggplot(aes(x = log(weight), y = n_offspring, colour = species))+
  geom_point(alpha = 0.5) +
  facet_wrap(~sex, ncol = 1) +
  theme_lemurs() +
  theme(legend.position = "none")
```


**R version, OS and packages used in this analysis:**
<!-- Leave this so people know the software they need to reproduce your work. -->
```{r}
sessionInfo()
```